---
title: "50.021 AI Project w CART, RF"
date: "12 June 2022"
author: "ESD + ISTD Trains"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: yes
    toc_depth: 2
---

# Clear Environment + Library
```{r, echo=TRUE,warning=FALSE,message=FALSE}
#use this to help you clear your environment withut clearing FUNCTIONS :)
rm(list = setdiff(ls(), lsf.str()))

## week 1-6
library(ggplot2)
library(psych)
library(ggfortify)
library(dplyr)
library(factoextra)
library(caTools)
library(ROCR)
library(mlogit)
library(leaps)
library(glmnet)

## week 8-12
library(rpart) # CART
library(rpart.plot) # CART plots
library(randomForest) # Random Forest
library(tm) # text mining; to process DTM
library(SnowballC) # Porter’s stemming algorithm
library(wordcloud) # Basic Visualisation with Wordcloud
library(e1071) # Naive Bayes Classifier

```

# Data Preparation

```{r}
# Load Train & Test Set
train <- read.csv("train_textcleaned.csv", stringsAsFactors=FALSE)
test <- read.csv("test_textcleaned.csv", stringsAsFactors=FALSE)

# Obtain Test Labels
testlabel <- test$label

# Extract relevant columns, insert placeholder test labels
train <- train[c(2,3)]
test <- test[c(2)]
test$label <- 5

# Combine the dataframes into one
twitter <- rbind(train,test)

#Take a look at proportion of tweets. Unequal proportion of sentiments. 5 means no sentiment
table(twitter$label)
table(testlabel)
```

# Pre-processing

To create the DTM, we will use the text mining package tm. Note that this is a development package, so there could be some variations based on the system you run (e.g., different R versions or different operating systems).

```{r}
corpus <- Corpus(VectorSource(twitter$tweet))
# corpus
```

Elements of the corpus.

```{r}
corpus[[1]]
as.character(corpus[[1]])

corpus[[4000]]
as.character(corpus[[4000]])
```

## Convert text to lower case

```{r}
corpus <- tm_map(corpus,content_transformer(tolower))
# corpus <- tm_map(corpus,tolower) # alternative command

as.character(corpus[[1]])
as.character(corpus[[4000]])
```

## Remove stopwords

Then, we move on to the stopwords. Here is the list of english stopwords in the `tm` package

```{r}
stopwords("english") # stopwords are common words found in english
```

Let's then remove the stopwords from the corpus

```{r}
corpus <- tm_map(corpus,removeWords,stopwords("english"))
# And ... Let's check a couple of documents
as.character(corpus[[1]])
as.character(corpus[[4000]])
```

## Remove punctuation

```{r}
corpus <- tm_map(corpus,removePunctuation)
# And ... Let's check a couple of documents
as.character(corpus[[1]])
as.character(corpus[[4000]])
```

## Stemming

Finally, we stem the words using Porter's stemming algorithm. Note that you may need to load the `SnowballC` package to use this functionality. The package is the R interface to the C libstemmer library that implements Porter's word stemming algorithm.

```{r}
corpus <- tm_map(corpus,stemDocument)
# And ... Let's check a couple of documents
as.character(corpus[[1]])
as.character(corpus[[4000]])
```

Let's compare the raw information with the post-processed one:

```{r}
twitter$tweet[1]
as.character(corpus[[1]])

twitter$tweet[4536]
as.character(corpus[[4536]])
```

## Create DTM

We can now create a document-term matrix from the original corpus.

```{r}
dtm <- DocumentTermMatrix(corpus)
dtm
```

Sparsity is close to 100%. (Most of them do not appear in the document)

Let's check the first & middle document

```{r}
dtm[1,]
inspect(dtm[1,])
```

```{r}
inspect(dtm[4000,])
```

## Removing sparse terms

An important information we can get is the frequency with which terms appear \
We can also check the frequency of specific words

```{r}
dtm[,"bitch"]
```

```{r}
dtm[,"fuck"]
```

In this specific case, we remove all terms with at least 99.5% empty entries

```{r}
dtm <- removeSparseTerms(dtm,0.995)
dtm # 
```

`dtm` is now a term-document matrix with 4536 documents and 282 terms.

# Preparing the DTM for model learning

We transform the term-document matrix into a matrix and, then, into a dataframe

```{r}
twittersparse <- as.data.frame(as.matrix(dtm))
# str(twittersparse)
# colnames(twittersparse)
```

This helps ensure that columns have valid names. For example, names starting with numbers are modified (e.g., 300k -> X300k).

```{r}
colnames(twittersparse) <- make.names(colnames(twittersparse))
# colnames(twittersparse)
```

# Train and Test a classifier

```{r}
# Split Train and Test Set
train1 <- subset(twittersparse[1:3175,])
test1 <- subset(twittersparse[3176:4536,])

# Last step, we add the output variable (polarity of tweets) to the `twittersparse` train1 dataframe. This is something we'll need to make predictions.

train1$label <- as.factor(twitter$label[1:3175])
# table(train1$label) # double-check numbers
```


## CART

```{r}
model2 <- rpart(label~., data=train1)
# summary(model2)
# model2

# Prediction, Confusion Matrix & Accuracy
predict2 <- predict(model2,newdata=test1, type = "class")

pred_table <- table(predict2,testlabel)
pred_table

accuracy <- (pred_table[1,1]+pred_table[2,2]+pred_table[3,3]+pred_table[4,4])/sum(pred_table)
accuracy 
```

## Random Forest

```{r}
# We use the default parameters to fit the model (500 trees)
model3 <- randomForest(label~.,data=train1, ntree=500)
# summary(model3)

# Prediction, Confusion Matrix & Accuracy
predict3 <- predict(model3,newdata=test1, type = "class")

pred_table <- table(predict3,testlabel)
pred_table

accuracy <- (pred_table[1,1]+pred_table[2,2]+pred_table[3,3]+pred_table[4,4])/sum(pred_table)
accuracy
```
